```{r, results="hide"}
library(data.table)
# install.packages("kableExtra")
library(knitr)
library(kableExtra)
```

# Estimating COVID-19 Case Severity in each CA County

Goal: Estimate what proportion of COVID-19 infections would require hospitalization in each US county based on reported COVID-19 severity by demographics and data on US county demographics.

---

## 1. Load Data

### Conditional Fatality Rates by Age

Using adjusted rates from https://www.imperial.ac.uk/media/imperial-college/medicine/sph/ide/gida-fellowships/Imperial-College-COVID19-NPI-modelling-16-03-2020.pdf instead of raw CCDC rates


```{r}
cond_rates = fread('data/imperial_rates.csv')

# transform rates to match earlier CCDC rates
cond_rates = cond_rates[,.(
  age_decade = age_group,
  hospitalizations_per_case = hospitalization_given_symptomatic,
  severe_cases_rate = (1-critical_care_given_hospitalization)*hospitalization_given_symptomatic,
  critical_case_rate = critical_care_given_hospitalization*hospitalization_given_symptomatic,
  case_fatality_rate = fatality_given_symptomatic
)]

cond_rates  %>% kable() %>% kable_styling()
```


### County Age Distribution

**Data source:**

Census County Population by Characteristics: 2010-2018 
(https://www.census.gov/data/datasets/time-series/demo/popest/2010s-counties-detail.html?)

Using 2018 population estimates. 


```{r}
age_buckets_by_county = fread('data/county_age_populations_acs.csv')
age_buckets_by_county[CTYNAME == "Santa Clara County"]  %>% kable() %>% kable_styling()
```



## 2. Estimating COVID-19 Case Severity in each County


---

**Steps to get simple estimate of severe cases:**

1. Join the county population numbers with conditional severity rates by age bucket.
2. Report the aggregated severity rates for each county.

---

#### Step 1: Join the county population numbers with CCDC conditional fatality rates by age bucket.
```{r}
merged_pop_fat_rate = merge(
  age_buckets_by_county[, .(FIPS, State = STNAME, County = CTYNAME, age_decade, population_in_age_group = pop)],
  cond_rates,
  by = 'age_decade')

merged_pop_fat_rate[County == "Santa Clara County"]  %>% kable(digits=4) %>% kable_styling()
```


#### Step 3: Report the aggregated severity rates for each CA county.

```{r}
county_case_severity_rates = merged_pop_fat_rate[,.(
  population = sum(population_in_age_group),
  hospitalizations_per_case = weighted.mean(hospitalizations_per_case, population_in_age_group),
  severe_cases_rate = weighted.mean(severe_cases_rate, population_in_age_group),
  critical_case_rate = weighted.mean(critical_case_rate, population_in_age_group),
  case_fatality_rate = weighted.mean(case_fatality_rate, population_in_age_group)
  ), by = c('FIPS', 'State', 'County')]

county_case_severity_rates[County == "Santa Clara County"] %>% kable(digits=3) %>% kable_styling()
```

```{r}
fwrite(county_case_severity_rates, 'data/county_case_severity_rates_v6.csv')
fwrite(merged_pop_fat_rate, 'data/county_age_severity_rates_v6.csv')

county_case_severity_rates[State == "California"][order(hospitalizations_per_case, decreasing = TRUE)] %>%
   kable(digits=3) %>% kable_styling()
```


